---
author: "Hugo Tameirão Seixas"
format: 
  html:
    embed-resources: true
editor_options: 
  chunk_output_type: console
params:
  area: "6"

---

```{r}
#| label: libraries
#| include: false

library(glue)
library(terra)
library(stars)
library(sf)
library(rstac)
library(gdalcubes)
library(janitor)
library(fs)
library(tidyverse)

```

```{r}
#| label: set-aoi
#| include: false

targets <- 
  read_sf("data/targets.gpkg") |>
  filter(id == params$area)

targets_area <- st_area(targets)

bbox <- st_bbox(targets)

bbox_sen <- targets |>
  st_transform("EPSG:3857") |>
  st_bbox()

```

```{r}
#| label: get-cube
#| include: false

if (!file_exists(glue("data/vi_cube_{params$area}.nc"))) {
  
  mpc_stac <- stac("https://planetarycomputer.microsoft.com/api/stac/v1/")

  sen_it <- mpc_stac %>%
    stac_search(
      collections = "sentinel-2-l2a",
      bbox = c(
        bbox["xmin"], bbox["ymin"],
        bbox["xmax"], bbox["ymax"]
      ),
      datetime = glue("{targets$start_date}/{targets$end_date}"),
      limit = 1000
    ) %>%
    get_request() %>%
    items_sign(
      sign_fn = sign_planetary_computer(
        headers = c(
          "Ocp-Apim-Subscription-Key" = "a0819286f5834452b5a3db24f5b19ecb"
        )
      )
    )
  
  sen_c <-
    stac_image_collection(
      s = sen_it$features,
      asset_names = c("B04", "B08"),
      property_filter = function(x) {x[["eo:cloud_cover"]] < 20}
    )

  sen_view <-
    cube_view(
      srs = st_crs(bbox_sen)$wkt,
      extent = list(
        t0 = as.character(targets$start_date), 
        t1 = as.character(targets$end_date),
        left = bbox_sen["xmin"], right = bbox_sen["xmax"],
        top = bbox_sen["ymax"], bottom = bbox_sen["ymin"]),
      dx = 10, dy = 10,
      dt = "P1M",
      aggregation = "median",
      resampling = "bilinear"
    )

  sen_cube <-
    raster_cube(
      image_collection = sen_c,
      view = sen_view
    ) %>%
    select_bands(c("B04","B08")) %>%
    write_ncdf(
      glue("data/vi_cube_{params$area}.nc"), 
      overwrite = TRUE
    )
  
}

```

```{r}
#| label: create-vis
#| include: false

gdalcubes_options(parallel = 16, ncdf_compression_level = 5)

sen_ts <- 
  ncdf_cube(glue("data/vi_cube_{params$area}.nc")) %>%
  apply_pixel("(B08-B04)/(B08+B04)", "NDVI") %>%
  extract_geom(targets) %>%
  as_tibble() %>%
  clean_names() %>%
  drop_na()

vis_ts <- sen_ts %>%
  mutate(
    time = ymd(time)
  ) %>%
  summarise(
    median = median(ndvi),
    min = quantile(ndvi, 0.10),
    max = quantile(ndvi, 0.90),
    .by = c(time, fid)
  ) %>%
  ggplot(
    aes(
      x = time, y = median,
      ymin = min, ymax = max,
      fill = fid, group = fid
    )
  ) +
  geom_line() +
  geom_ribbon(alpha = 0.5, fill = "#c4c4c4") +
  scale_x_date(
    date_labels = "%F", 
    date_breaks = "12 months",
    expand = c(0.01, 0.01)
  ) +
  labs(y = "NDVI") +
  theme_minimal() +
  theme(
    axis.title.x = element_blank()
  )

sen_stars <- 
  ncdf_cube(glue("data/vi_cube_{params$area}.nc")) %>%
  apply_pixel("(B08-B04)/(B08+B04)", "NDVI") %>%
  reduce_time("median(NDVI)") %>%
  filter_geom(st_geometry(targets)) %>%
  st_as_stars()

vis_map <- ggplot() +
  geom_stars(data = sen_stars) +
  geom_sf(
    data = st_transform(targets, "EPSG:3857"), 
    fill = NA,
    linewidth = 2,
    color = "#000000"
  ) +
  labs(fill = "NDVI") +
  scale_fill_distiller(
    palette = "YlOrBr", 
    na.value = "transparent",
    direction = 0
  ) +
  theme_void()

vis_hist <- sen_stars |>
  as_tibble() |>
  ggplot() +
  geom_histogram(
    aes(x = NDVI_median, fill = after_stat(x)),
    color = "#000000"
  ) +
  labs(x = "NDVI", y = "Count") +
  scale_fill_distiller(
    palette = "YlOrBr", 
    na.value = "transparent",
    direction = 0
  ) +
  theme_minimal() +
  theme(
    legend.position = ""
  )

```

# Analysis of area `r params$area`

:::: {.columns}

::: {.column width="45%"}

This is an agriculture field. 

NDVI was calculated using Sentinel 2 images. 

It covers an area of **`r round(targets_area * 0.0001)` ha**.

The map shows the median values from **`r min(sen_ts$time)`** to **`r max(sen_ts$time)`**.

```{r}
#| label: plot-hist
#| echo: false
#| warning: false

vis_hist

```

:::

::: {.column width="10%"}
<!-- empty column to create gap -->
:::

::: {.column width="45%"}

```{r}
#| label: plot-map
#| echo: false

vis_map

```

It is possible to observe the variation of the median over the area. This can indicate areas with different vegetation cover along the years.

When analyzing the historical median NDVI, we can observe that the maximum value is **`r round(max(sen_stars$NDVI_median, na.rm = T), digits = 2)`**, the minimum value is **`r round(min(sen_stars$NDVI_median, na.rm = T), digits = 2)`**, and the median value is **`r round(median(sen_stars$NDVI_median, na.rm = T), digits = 2)`**

:::

::::

```{r}
#| label: plot-ts
#| echo: false
#| fig-asp: 0.2
#| out-width: 100%

vis_ts

```

We can also observe the temporal variation of the NDVI in this field.
